---
title: "Confidence Intervals"
output: slidy_presentation
css: "www/teachingnotes.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The Goal

You've asked how to build this type of visualisation:

```{r, message=FALSE, warning=FALSE}
library(plotly)
x <- rnorm(100)
y <-  + .7*x + rnorm(100)
f1 <- as.factor(c(rep("A",50),rep("B",50)))
f2 <- as.factor(rep(c(rep("C",25),rep("D",25)),2))
df <- data.frame(cbind(x,y))
df$f1 <- f1
df$f2 <- f2
{ggplot(df,aes(x=x,y=y)) +
    geom_point() +
    facet_grid(f1~f2) +
    stat_smooth(method="lm") }%>%
  ggplotly()
```

But first we need to be able to build this chart:

```{r}
library(ggplot2)
x <- rnorm(100)
y <-  + .7*x + rnorm(100)
f1 <- as.factor(c(rep("A",50),rep("B",50)))
f2 <- as.factor(rep(c(rep("C",25),rep("D",25)),2))
df <- data.frame(cbind(x,y))
df$f1 <- f1
df$f2 <- f2
ggplot(df,aes(x=x,y=y)) +
    geom_point() +
    stat_smooth(method="lm")
```

# Tools/Techniques we need

- Importing Data
- Setting Factors
- ggplot2
- facets in ggplot2
- Making ggplot2 interactive with ggplotly

# Importing Data

```{r,echo=F,eval=F}
df <- data.frame(
  "metric1" = rnorm(100),
  "metric2" = + .7*x + rnorm(100),
  "category" = c(rep("Rural",50),rep("Urban",50)),
  "country" = rep(c(rep("India",25),rep("Vietnam",25)),2)
)
library(xlsx)
write.xlsx2(file = "data/confidence-data.xlsx", x = df,row.names = F)
```

The data for this example is stored within an .xlsx file, which can be imported directly using the `xlsx` library. However:

<div class="incremental">
- The `xlsx` library depends on a specific version of Java being installed and set as an appropriate PATH variable.

- It is tremendously slow.

```{r, echo=T}
library(xlsx)
confidence_data <- read.xlsx2("data/confidence-data.xlsx", sheetIndex = 1)
```

- It's also quite dumb

```{r}
str(confidence_data)
```

</div>

# xlsx_convert_import

My personal recommendation is to import as .xlsx and export to csv, it makes data processing much simpler. This utility function provides a simple way to achieve this:

```{r, echo=T}
xlsx_convert_import <- function(inputFile = NA, outputFile = NA){
  if(file.exists(outputFile)){
    imported_data <<- read_csv(outputFile)
  } else {
    library(xlsx)
    xlsx_import <- read.xlsx2(inputFile, sheetIndex = 1)
    write.csv(xlsx_import, file = outputFile, row.names = FALSE)
    imported_data <<- read_csv(outputFile)
  }
}
```

Note that the function uses the `read_csv` function from `readr`:

```{r}
library(readr)
confidence_data <- xlsx_convert_import(inputFile = "data/confidence-data.xlsx", outputFile = "data/confidence-data.csv")
str(confidence_data)
```

# ggplot2

The `ggplot2` provides us a complete "grammar of graphics", a consistent methodology for building and adding to charts (as well as maps and network diagrams!).

Documentation for `ggplot2` is available here [http://docs.ggplot2.org/current/](http://docs.ggplot2.org/current/) and there is also a great cheat sheet available under Help in the menubar.

# ggplot and aes

The workhorses of a `ggplot2` plot are as follows:

- `ggplot(data = df, ...)` initiates a ggplot object with `df` as the `data` to be used within the object.

- `aes` specifies the "aesthetic mappings" for how variables in the `data` are mapped to visual properties of the object

We can therefore create an empty chart as follows:

```{r, echo=T}
ggplot(confidence_data ,aes(x = metric1, y = metric2))
```

# Adding layers

To include additional visual elements to the chart, add "layers" with the `+` operator:

```{r, echo=T}
ggplot(confidence_data ,aes(x = metric1, y = metric2)) +
  ggtitle("Regressing metric 1 on metric 2")
```

Often users choose to make intermediatry assignments in building their charts:

```{r, echo=T}
regress_chart <- ggplot(confidence_data ,aes(x = metric1, y = metric2)) +
  ggtitle("Regressing metric 1 on metric 2")


regress_chart + xlab("some metric")
```


# geoms

Geoms are used to add visual features to a chart, using the `aes` mappings between data variables and the chart's coordinate system:

```{r, echo=T}
regress_chart <- ggplot(confidence_data ,aes(x = metric1, y = metric2)) +
  ggtitle("Regressing metric 1 on metric 2")

regress_chart + geom_point()
```

Regression and associated confidence levels can easily be plotted together using the `stat_smooth` function:

```{r}
regress_chart + geom_point() + stat_smooth(method="lm")
```

# Where is that model?

The models used by `stat_smooth` are accessible directly using `loess` (the automatically chosen fitting method for small datasets)

```{r}
confidence_loess_model <- loess(metric2 ~ metric1, confidence_data)
```

Models have many properties, as with most objects these can be listed with `names`

```{r}
names(confidence_loess_model) 
```

What about real models? loess is for exploring relations

# lm

All models use the same syntax, so to compute a linear model use:

```{r}
confidence_lm_model <- lm(metric2 ~ metric1, confidence_data)
```

Frustratingly, the properties of the model that you might assume to live in the object aren't there! Where is the RSquared?

```{r}
names(confidence_lm_model)
```


<div class="incremental">

To generate the R squared value, and other useful properties, you must use `summary` and extract from that object

```{r}
lm_model_summary <- summary(confidence_lm_model)
lm_model_summary$adj.r.squared
```
</div>

# Back to ggplot and plotly

The `plotly` library allows us to convert "simple" `ggplot` objects directly into interactive charts:

```{r}
library(plotly)
regress_chart <- ggplot(confidence_data ,aes(x = metric1, y = metric2)) +
  ggtitle("Regressing metric 1 on metric 2")

ggplotly(regress_chart + geom_point() + stat_smooth(method="lm"))
```

# Faceting charts

The dataset we're looking at has a number of different facets, it's useful to be able to split our chart by these.

In order to do that, we need to provide `ggplot` with a `factor` variable:

```{r}
confidence_data$country <- as.factor(confidence_data$country)
confidence_data$category <- as.factor(confidence_data$category)
```

<div class="incremental">

`facet_grid` allows a chart to be split according to a factor(s):

- `facet_grid(. ~country)` will facet only on country by column

```{r}
regress_chart + facet_grid(. ~country) + geom_point() + stat_smooth(method="lm")
```

- `facet_grid(. ~country)` will facet on country by row

```{r}
regress_chart + facet_grid(country ~ .) + geom_point() + stat_smooth(method="lm")
```

- `facet_grid(country ~ category)` will facet on country and country

```{r}
regress_chart + facet_grid(country ~ category) + geom_point() + stat_smooth(method="lm")
```

- `ggplotly` will happily gobble up all of these

```{r}
ggplotly(regress_chart + facet_grid(country ~ category) + geom_point() + stat_smooth(method="lm"))
```

</div>
