---
title: "Stata Import"
output: slidy_presentation
runtime: shiny
css: "www/teachingnotes.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing Data in RStudio

Importing data in RStudio is much easier if you work in projects, which is discussed in the main body of the training materials (/Development/Lecturer-Notes/RMarkdown.Rmd).

This is because projects set the working directory to the directory containing the .RProj file, and all paths can be made relevant to that.

In this project folder there is a directory called data, which may be accessed using:

```{r, eval=FALSE}
list.files("data")
```

# .dta files

There are two libraries for importing .dta libraries:

- `library(foreign)` can only be used for importing .dta from Stata v12 and earlier

- `library(readstata13)` can be used to import current generation .dta files

# Example File

There are a multitude of example data files here: http://www.stata-press.com/data/r14.

For quickness I tried to get the Iris fish dataset but ended up with a different fishy dataset, this can easily be imported as follows:

```{r}
library(readstata13)
fish_dta <- read.dta13("data/fish.dta")
```

To better understand this data structure, use `class` and `str`

```{r,echo=FALSE, eval=F}
str(fish_dta)
```

# What is that thing?

The `read.dta13` function imports the .dta file as a `data.frame` with attributes. 

Most `data.frame` don't have attributes, because they only contain data; but .dta files also contain information about the dataset.

- Accessing the data is the same as for all `data.frame`
- The attributes are accessible through dedicated functions.

# Attributes

Return all of an object's attributes with `attributes`:

```{r, eval=F}
attributes(fish_dta)
```


Individual attributes can be accessed as follows:

```{r, eval=F}
attr(fish_dta, "formats")
```


Programmatically extract all attribute names as follows:

```{r, eval=F}
names(attributes(fish_dta))
```

# Data.frames

There are two approaches to extracting data from data.frames:

- Base R: Extract columns and rows directly from the `data.frame` object

<div class="incremental">
```{r, eval = F}
fish_dta$
```

```{r, eval = F}
fish_dta[1, ]
```

```{r, eval = F}
fish_dta[, 1]
```
</div>

- dplyr: `dplyr` is the data filtering and processing library from tidyverse, and makes many operations simpler than they would be using base R 

<div class="incremental"> 

```{r, eval = F}
fish_dta %>%
  select(camper, persons)
```

```{r, eval = F}
fish_dta %>%
  select(-camper)
```

</div>

# Filtering Data

Filtering data with `dplyr` is very simple:

```{r, eval = F}
fish_dta %>%
  select(-nofish) %>%
  filter(xb > -0.8 &
           persons == 1)
```

# How about extracting columns by strings?

In shiny it is important to remember that `input$variables` are typically strings, it is therefore necessary to be able to use `dplyr` with strings for column names.

This is simply achieved by appending a `_` to every function you use:

```{r, eval = F}
fish_dta %>%
  select_("-nofish") %>%
  filter_("xb" > -0.8 &
           "persons" == 1)
```

The "default" `dplyr` functions rely on non-standard evaluation to access column names, if you wish to read more about this do watch out for dragons.
